### 2016-11-05 conclusions.tex
saturday more shit

\section{Conclusions \& Future Work}
%\subsection{Conclusions}
Pruning neurons (instead of pruning individual weights) in a pre-trained neural network without seeing a major loss in performance is not only possible but also enables compressing networks to 40-80\% of their original size, which is of great importance in constrained memory environments like embedded devices. This fact is established through the experiments using the brute force criterion, which if made computationally viable (through parallelization) or approximated more efficiently than the Taylor Series based methods discussed in this paper, can prove to be a useful compression tool. It would also be interesting to see how these methods perform on deeper networks and on some other popular and real world datasets. Also, as mentioned in Section \ref{sec2}, we have not considered all possible combinations of neuron interdependence in this work due to the in-feasibility of implementation and have always pruned one neuron at a time. Even though the greedy evaluation of such combinations is highly prohibitive, it is not entirely implausible to think that algorithms can be developed for it in the future. That would help in truly tapping into the power of these interconnections and hopefully lead to impressive performance results.

>>>>

\section{Conclusions \& Future Work}
%\subsection{Conclusions}
Pruning neurons (instead of pruning individual weights) in a pre-trained neural network without seeing a major loss in performance is not only possible but also enables compressing networks to 40-70\% of their original size, which is of great importance in constrained memory environments like embedded devices. This fact is established through the experiments using the brute force criterion, which if made computationally viable (through parallelization) or approximated more efficiently than the Taylor Series based methods discussed in this paper, can prove to be a useful compression tool. It would also be interesting to see how these methods perform on deeper networks and on some other popular and real world datasets. Also, as mentioned in Section \ref{sec2}, we have not considered all possible combinations of neuron interdependence in this work due to the in-feasibility of implementation and have always pruned one neuron at a time. Even though the greedy evaluation of such combinations is highly prohibitive, it is not entirely implausible to think that algorithms can be developed for it in the future. That would help in truly tapping into the power of these interconnections and hopefully lead to impressive performance results.

### 2016-11-05 introduction.tex
saturday more shit

Essentially, \cite{mozer1989using} observed that neural networks during training do \textit{not} distribute the learning representation evenly or equitably across hidden units. What actually happens is that a few, elite neurons learn an approximation of the input-output function, and the remaining units must learn a complex interdependence function which cancels out their respective influence on the network output. Furthermore, assuming enough units exist to learn the function in question, increasing the number of parameters does not increase the richness or robustness of the learned approximation, but rather simply increases the likelihood of overfitting and the number of noisy parameters to be canceled during training. This is evinced by the fact that in many cases, multiple neurons can be removed from a network with no re-training and with negligible impact on the quality of the output approximation. In other words, there are few bipartisan units in a trained network. A unit is typically either part of the (possibly overfit) input-output function approximation, or it is part of an elaborate noise cancellation task force. Assuming this is the case, most of the compute-time spent training a neural network is likely occupied by this arguably wasteful procedure of silencing superfluous parameters, and pruning can be viewed as a necessary treatment procedure to ``trim the fat.''

@@

It is well known that too many free parameters in a neural network can lead to overfitting. Regardless of the number of weights used in a given network, as \cite{segee1991fault} assert, the representation of a learned function approximation is almost never evenly distributed over the hidden units, and thus the removal of any single hidden unit at random can actually result in a network fault. \cite{mozer1989using} argue that only a subset of the hidden units in a neural network actually latch on to the invariant or generalizing properties of the training inputs, and the rest learn to either mutually cancel each other's influence or begin over-fitting to the noise in the data. We leverage this idea in the current work to rank all neurons in pre-trained networks based on their effective contributions to the overall performance. We then remove the unnecessary neurons to reduce the network's footprint. Through our experiments we not only concretely validate the theory put forth by \cite{mozer1989using} but we also successfully build on  it to prune networks to 40 to 60 \% of their original size without any major loss in performance.

>>>>

Essentially, \cite{mozer1989using} observed that neural networks during training do \textit{not} distribute the learning representation evenly or equitably across hidden units. What actually happens is that a few, elite neurons learn an approximation of the input-output function, and the remaining units must learn a complex interdependence function which cancels out their respective influence on the network output. Furthermore, assuming enough units exist to learn the function in question, increasing the number of parameters does not increase the richness or robustness of the learned approximation, but rather simply increases the likelihood of overfitting and the number of noisy parameters to be canceled during training. This is evinced by the fact that in many cases, multiple neurons can be removed from a network with no re-training and with negligible impact on the quality of the output approximation. In other words, there are few bipartisan units in a trained network. A unit is typically either part of the (possibly overfit) input-output function approximation, or it is part of an elaborate noise cancellation task force. Assuming this is the case, most of the compute-time spent training a neural network is likely occupied by this arguably wasteful procedure of silencing superfluous parameters, and pruning can be viewed as a necessary procedure to ``trim the fat.''

@@

It is well known that too many free parameters in a neural network can lead to overfitting. Regardless of the number of weights used in a given network, as \cite{segee1991fault} assert, the representation of a learned function approximation is almost never evenly distributed over the hidden units, and thus the removal of any single hidden unit at random can actually result in a network fault. \cite{mozer1989using} argue that only a subset of the hidden units in a neural network actually latch on to the invariant or generalizing properties of the training inputs, and the rest learn to either mutually cancel each other's influence or begin overfitting to the noise in the data. We leverage this idea in the current work to rank all neurons in pre-trained networks based on their effective contributions to the overall performance. We then remove the unnecessary neurons to reduce the network's footprint. Through our experiments we not only concretely validate the theory put forth by \cite{mozer1989using} but we also successfully build on  it to prune networks to 40 to 60 \% of their original size without any major loss in performance.

### 2016-11-05 results.tex
saturday more shit

\section{Experimental Results}
For all the results presented in this section, the MNIST database of Handwritten Digits by \cite{lecun-mnisthandwrittendigit-2010} was used. It is worth noting that due to the time taken by the brute force algorithm we rather used a 5000 image subset of the MNIST database in which we have normalized the pixel values between 0 and 1.0, and compressed the image sizes to 20x20 images rather than 28x28, so the starting test accuracy reported here appears higher than those reported by LeCun et al. We do not believe that this affects the interpretation of the presented results because the basic learning problem does not change with a larger dataset or input dimension.

>>>>

\section{Experimental Results}
\subsection{Example Regression Problem}
This problem serves as a quick example to demonstrate many of the phenomena described in previous sections. We trained two networks to learn the cosine function, with one input and one output. This is a task which requires no more than 11 sigmoid neurons to solve entirely, and in this case we don't care about overfitting because the cosine function has a precise definition. Furthermore, the cosine function is a good toy example because it is a smooth continuous function and, as demonstrated by \cite{nielsen2015neural}, if we were to tinker directly with the weights and bias parameters of the network, we could allocate individual units within the network to be responsible for constrained ranges of inputs, similar to a basis spline function with many control points. This would distribute the learned function approximation evenly across all hidden units, and thus we have presented the network with a problem in which it could productively use as many hidden units as we give it. In this case, a pruning algorithm would observe a fairly consistent increase in error after the removal of each successive unit. In practice however, regardless of the number of experimental trials, this is not what happens. The network will always use 10-11 hidden units and leave the rest to cancel each other's influence.

\begin{figure}[!h]
\centering
\includegraphics[width=0.49\linewidth]{png/cos-small.pdf}
\includegraphics[width=0.49\linewidth]{png/cos-big.pdf}
\caption{Degradation in squared error after pruning a two-layer network trained to compute the cosine function (\textbf{Left Network:} 2 layers, 10 neurons each, 1 output, logistic sigmoid activation, starting test accuracy: 0.9999993, \textbf{Right Network:} 2 layers, 50 neurons each, 1 output, logistic sigmoid activation, starting test accuracy: 0.9999996)}
\label{fig:cosine-double-layer}
\end{figure}

Figure \ref{fig:cosine-double-layer} shows two graphs. Both graphs demonstrate the use of the iterative re-ranking algorithm and the comparative performance of the brute-force pruning method (in blue), the first order method (in green), and the second order method (in red). The graph on the left shows the performance of these algorithms starting from a network with two layers of 10 neurons (20 total), and the graph on the right shows a network with two layers of 50 neurons (100 total).

In the left graph, we see that the brute-force method shows a graceful degradation, and the error only begins to rise sharply after 50\% of the total neurons have been removed. The error is basically constant up to that point. In the first and second order methods, we see evidence of poor decision making in the sense that both made mistakes early on, which disrupted the output function approximation. The first order method made a large error early on, though we see after a few more neurons were removed this error was corrected somewhat (though it only got worse from there). This is direct evidence of the lack of fault tolerance in a trained neural network. This phenomenon is even more starkly demonstrated in the second order method. After making a few poor neuron removal decisions in a row, the error signal rose sharply, and then went back to zero after the 6th neuron was removed. This is due to the fact that the neurons it chose to remove were trained to cancel each others' influence within a localized part of the network. After the entire group was eliminated, the approximation returned to normal. This can only happen if the output function approximation is not evenly distributed over the hidden units in a trained network.

This phenomenon is even more starkly demonstrated in the graph on the right. Here we see the first order method got ``lucky'' in the beginning and made decent decisions up to about the 40th removed neuron. The second order method had a small error in the beginning which it recovered from gracefully and proceeded to pass the 50 neuron point before finally beginning to unravel. The brute force method, in sharp contrast, shows little to no increase in error at all until 90\% of the neurons in the network have been obliterated. Clearly first and second order methods have some value in that they do not make completely arbitrary choices, but the brute force method is far better at this task.

This also demonstrates the sharp dualism in neuron roles within a trained network. These networks were trained to near-perfect precision and each pruning method was applied \textit{without} any re-training of any kind. Clearly, in the case of the brute force or oracle method, up to 90\% of the network can be completely extirpated before the output approximation even begins to show any signs of degradation. This would be impossible if the learning representation were evenly or equitably distributed. Note, for example, that the degradation point in both cases is approximately the same. This example is not a real-world application of course, but it brings into very clear focus the kind of phenomena we will discuss in the following sections.

\subsection{Results on MNIST Dataset}
For all the results presented in this section, the MNIST database of Handwritten Digits by \cite{lecun-mnisthandwrittendigit-2010} was used. It is worth noting that due to the time taken by the brute force algorithm we rather used a 5000 image subset of the MNIST database in which we have normalized the pixel values between 0 and 1.0, and compressed the image sizes to 20x20 images rather than 28x28, so the starting test accuracy reported here appears higher than those reported by LeCun et al. We do not believe that this affects the interpretation of the presented results because the basic learning problem does not change with a larger dataset or input dimension.
